Study of the hadronization of charged pions

### 2019-03-14 chap5.tex
Edits of first part of chap5

Pions are detected in the angular range from $\sim$10 to $\sim$140 degrees 
using only DC and SC signals. To identify negatively charged pions, we select 
negative tracks which were not identified as electrons already.

@@

The identification of positively charged pions is similar to that of negative 
pions. However, the time of flight plot is significantly more busy (see figure 
\ref{PipTOF}), due to an important contamination from $K^+$s and protons at a
high momentum. As the CC is not efficient enough for a hadron separation, the 
numerous kaons and protons should be removed by the following tighter TOF cuts:

@@

The much more apparent contamination sources than for negative pions, motivates us to study 
the TOF profile much more thoroughly. One can see in figures~\ref{TOF-1}, \ref{TOF-2} and 
\ref{TOF-3} the $\Delta \beta$ profiles for many momentum slices. In figure~\ref{TOF-1}
one can see some contamination from lighter leptons ($e+$ and $\mu^+$). These represent
a very small amount and one can see some justification for the choice of a $\pm 0.03$
cut as these particles do not appear to be discernable beyond that level. Similarly
in figure~\ref{TOF-2} one can see the contribution from kaons. The larger contribution
of this contamination motivated the tightening of the TOF cut on this side, to the limit
where kaons are a minority in the tail of the pion distribution. Finally a similar 
method is applied for the even larger proton contamination as shown in figure \ref{TOF-3}.

@@

These cuts minimize the kaon contamination, below 2.5~GeV/c, and the 
proton contamination for all momenta. The kaon contribution cannot be avoided, however, it should remain small ($\sim$ 3\% according to simulation) 
with only a small effect on our final results (see section \ref{SysId} 
for more details). Because protons could lead to even more contamination than kaons, 
a stricter cut is used at a high momentum. This cut is also justified by 
HERMES data \cite{Airapetian:2007vu}, which showed a very different behavior of protons compared to other hadrons.

@@

To differentiate between the two targets and remove the background, we need to 
determine the origin of our final-state particles. As illustrated in figure~\ref{vertex}, the small misalignment of the beam with the CLAS detector led to an artificial shift on different sectors' z-vertex positions. To correct from this effect, we added the shift shown in table~\ref{tab:vertex} when determining each particle's vertex position. Those values were obtained from the fit of the solid target distribution on each sector.

@@

The position of our targets could also vary from a run to another, but this was not significant in this experiment. Indeed, except for aluminum, other targets remained in the same position within one or two millimeters. All targets' positions, in CLAS coordinates, are given in table \ref{tab:targets}. 
We show in figures \ref{VertexSolid} and \ref{VertexLiquid} the evolution 
of the position of the targets with run number. The data where the aluminum
target is off by 1 cm has been excluded from the analysis. Other variations
appear to be of a mm or less, while the gap between the solid and liquid target
is about 5 cm. They should therefore have negligible contribution to the 
error on acceptance corrections in regard to the error budget already associated
to the procedure.

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=7.5cm] {answer-fig/VertexC.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexAl.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexFe.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexSn.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexPb.png} 
\caption {Vertex origin along $z$ (in cm) of electrons as a function of run 
number for the different solid targets.}
\label{VertexSolid}
\end{figure}

\begin{figure}[tbp]
\centering
\includegraphics[width=7.5cm] {answer-fig/VertexDeutC.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutFe.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutSn.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutPb.png} 
\caption {Vertex origin along $z$ (in cm) of electrons as a function of run 
number for the deuterium target in different solid targets configurations.}
\label{VertexLiquid}
\end{figure}

@@

To check the quality of the runs\footnote{Runs are typically stopped every two hours
of data taking, but they can be smaller when experimental problems arrise.}, we 
monitored the ratio of scattered electrons yields obtained from solid and liquid 
targets. In case of the beam hitting other materials in the beam-line or of any 
other problems that would affect our final result, 
this ratio will be off and indicate the problematic runs. In figure 
\ref{DataQ}, we show the electron yield ratios for different nuclear target; we 
fitted these ratios to determine the mean value for each target, then we discarded 
runs that were 5~$\sigma$ away from the appropriate mean. Must note that these 
ratios were consistent with our target's thicknesses given in \cite{Hakobyan:2008zz} 
except for a carbon target. For that reason, the density of the latter was 
recently remeasured, and its new value was finally matching our 
data\footnote{i.e. (1.747+/-0.0007)g/cm$^3$ instead of 2.235 g/cm$^3$}.

@@

@@

We also apply a cut on the energy transferred fraction, $y = \nu/E_{beam} < 0.85$. This cut aims to remove the low energy electrons that are significantly affected by radiative effects and sat at the limit of our trigger threshold.

@@

Following the first round of questions from the analysis review committee, we
reflected about this implementation of error bars and possible overlap with the 
systematic errors evaluated below. First we found an error with how the formula
above was implemented and corrected it. This reduces the size of statistical 
error bars of the transverse momentum broadening by roughly a factor 2 to 3.

TODO See new results and re evaluate statement below.
We found first that the correlation between the weights should reduce
the statistical errors. In a given bin the weights associated with each 
targets are always of similar sizes, this correlation will reduce the size of 
error bars. However, we do not see a method to properly take this 
effect fully into account in term of purely statistical error bars. In order
to solve this issue we came back to a point by the committee on the topic of 
systematic error bars (Q 24). There it was noted that some of these effects are
double counted once in the statistical error and once in the systematic. As these
errors are solely due to the acceptance correction, we decided to account them 
only in the systematic error budget. So we reverted statistical errors to equations
\ref{eq:MRError} and \ref{eq:DPtError} and keep the full systematic errors as they
were evaluated before.

>>>>

Pions are detected starting at a momentum of 250 MeV in the angular range 
from $\sim$10 to $\sim$140 degrees 
using only DC and SC signals. To identify negatively charged pions, we select 
negative tracks which were not identified as electrons already.

@@

The identification of positively charged pions is similar to that of negative 
pions. However, the time of flight plot is significantly more busy (see figure 
\ref{PipTOF}), due to an important contamination from $K^+$s and protons at a
high momentum. As the CC is not efficient enough for a hadron separation, the 
kaons and protons contributions are reduced by the following tighter TOF cuts:

@@

In opposition with negative pions, contamination sources are apparent and motivate a careful 
study of the TOF profile. One can see in figures~\ref{TOF-1}, \ref{TOF-2} and 
\ref{TOF-3} the $\Delta \beta$ profiles for many momentum slices. In figure~\ref{TOF-1}
one can see some contamination from lighter leptons ($e^+$ and $\mu^+$). These represent
a very small amount and one can see some justification for the choice of a $\pm 0.03$
cut as these particles do not appear to be discernable beyond that level. Similarly
in figure~\ref{TOF-2} one can see the contribution from kaons. The larger contribution
of this contamination motivated the tightening of the TOF cut on this side, to the limit
where kaons are a minority in the tail of the pion distribution. Finally a similar 
method is applied for the even larger proton contamination as shown in figure \ref{TOF-3}.

@@

These cuts minimize the kaon contamination, below 2.5~GeV/c, and the 
proton contamination for all momenta. The kaon contribution cannot be avoided, however, it should remain small (peaking $\sim$ 3\% according to our simulation) 
with only a small effect on our final results (see section \ref{SysId} 
for more details). Because protons could lead to even more contamination than kaons, 
a stricter cut is used at a high momentum. This cut is also justified by 
HERMES data \cite{Airapetian:2007vu}, which showed a very different behavior of protons compared to other hadrons.

@@

To differentiate between the two targets and remove the background, we need to 
determine the point of origin of our final-state particles. As illustrated in figure~\ref{vertex}, the small misalignment of the beam with the CLAS detector led to an artificial shift on different sectors' z-vertex positions. To correct from this effect, we added the shift shown in table~\ref{tab:vertex} when determining each particle's vertex position. Those values were obtained from the fit of the solid target distribution on each sector.

@@

The position of our targets could also vary from a run to another, but this was not significant in this experiment. Indeed, except for aluminum, other targets remained in the same position within one or two millimeters. All targets' positions, in CLAS coordinates, are given in table \ref{tab:targets}. 
We show in figures \ref{VertexSolid} and \ref{VertexLiquid} the evolution 
of the position of the targets with run number. The data where the aluminum
target is off by 1 cm has been excluded from the analysis. Other variations
appear to be of a mm or less, while the gap between the solid and liquid target
is about 5 cm. They should therefore have negligible contribution to the 
error on acceptance corrections in regard to the error budget already associated
to the measurements.

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=7.5cm] {answer-fig/VertexC.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexAl.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexFe.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexSn.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexPb.png} 
\caption {Vertex origin along $z$ (in cm) of electrons as a function of run 
number for the different solid targets. Top left is for carbon, top right
aluminum, middle left iron, middle right tin and bottom is for lead.}
\label{VertexSolid}
\end{figure}

\begin{figure}[tbp]
\centering
\includegraphics[width=7.5cm] {answer-fig/VertexDeutC.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutFe.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutSn.png} 
\includegraphics[width=7.5cm] {answer-fig/VertexDeutPb.png} 
\caption {Vertex origin along $z$ (in cm) of electrons as a function of run 
number for the deuterium target in different solid targets configurations.
Top left is for carbon, top right iron, bottom left tin and bottom right is for lead.}
\label{VertexLiquid}
\end{figure}

@@

To check the quality of the runs\footnote{Runs are typically stopped every two hours
of data taking, but they can be smaller when experimental problems arrise.}, we 
monitored the ratio of scattered electrons yields obtained from solid and liquid 
targets. In case of the beam hitting other materials in the beam-line or of any 
other problems that would affect our final result, 
this ratio will be off and indicate the problematic runs. In figure 
\ref{DataQ}, we show the electron yield ratios for different nuclear target; we 
fitted these ratios to determine the mean value for each target, then we discarded 
runs that were 5~$\sigma$ away from the appropriate mean. Must note that these 
ratios were consistent with our target's thicknesses given in \cite{Hakobyan:2008zz} 
except for the carbon target. For that reason, the density of the latter was 
remeasured, and its new value was finally matching our 
data\footnote{i.e. (1.747+/-0.0007)g/cm$^3$ instead of 2.235 g/cm$^3$}.

@@

%TODO provide an updated target table

@@

We also apply a cut on the energy transferred fraction, $y = \nu/E_{beam} < 0.85$. This cut aims to remove the low energy electrons that are significantly affected by radiative effects and are at the limit of our trigger threshold.

@@

Following the first round of questions from the analysis review committee, we
reflected about this implementation of error bars and possible overlap with the 
systematic errors evaluated below. First we found an error with how the formula
for transverse momentum broadening was implemented and corrected it. This 
reduces the size of statistical 
error bars of the transverse momentum broadening by roughly a factor 2 to 3. 
We also reflected that the correlation between the weights should reduce
the statistical errors. In a given bin the weights associated with each 
targets are always of similar sizes, this correlation will reduce the size of 
error bars. However, we do not see a method to properly take this 
effect fully into account in term of purely statistical error bars, and
remains as is. Finally, 
it was noted that some of the effects were double counted once through the 
statistical error on the acceptance correction and the systematics associated
with the correction. So, we decided to account only the systematic error budget
and use purely statistical errors of equations
\ref{eq:MRError} and \ref{eq:DPtError}. In summary, we concluded that the 
systematic error contribution associated to the statistics of the 
simulation sample were indeed a double counting and are not used anymore
in any error bars.
